### 2017-12-21 inst/rmd/vis_sim.Rmd
Add function for visualizing simulation results

---
title: "Simulation Explorer"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r global, include=FALSE}

library(plotly)
library(viridis)
library(shiny)
e <- new.env() # make a new environment to keep the global clean
e$sim <- temp  # temp object (in global) created by vis_sim function
rm(temp)       # clean-up
e$ages <- e$sim$ages
e$years <- e$sim$years
e$vis_distribution <- "sp_N" %in% names(e$sim)

```

Abundance
================================================================================

Row
--------------------------------------------------------------------------------

### Total abundance

```{r}
e$tot <- data.frame(Year = e$years, N = colSums(e$sim$N))
e$tot$text <- paste0("Year: ", e$tot$year, "\nN: ", round(e$tot$N))
plot_ly(x = ~Year, y = ~N, text = ~text, hoverinfo = "text",
        type = "scatter", mode = "lines", data = e$tot) %>%
  layout(xaxis = list(rangeslider = list(type = "date")))
```

Row
--------------------------------------------------------------------------------

### Abundance at age

```{r}
e$d <- as.data.frame.table(e$sim$N, responseName = "N")
names(e$d) <- c("Age", "Year", "N")
e$d$text <- paste0("Age: ", e$d$Age, "\nYear: ", e$d$Year, "\nN: ", round(e$d$N))
plot_ly(x = ~Age, y = ~Year, z = ~N, color = ~Age,
        text = ~text, hoverinfo = "text",
        colors = viridis(length(e$ages)),
        type = "scatter3d", mode = "lines",
        line = list(width = 8),
        data = e$d)
```

### Total mortality

```{r}
plot_ly(x = e$ages, y = e$years, z = e$sim$Z, type = "surface") %>% 
  layout(
    scene = list(
      xaxis = list(title = "Age"),
      yaxis = list(title = "Year"),
      zaxis = list(title = "Z")
    ))
```

Distribution { `r if(!e$vis_distribution) ".hidden"`}
================================================================================

Column {.sidebar}
--------------------------------------------------------------------------------

```{r}

sliderInput("age", "Age", min = min(e$ages), max = max(e$ages), 
            value = 1, step = 1)
sliderInput("year", "Year", min = min(e$years), max = max(e$years), 
            value = 1, step = 1)
selectInput("type", "Type", choices = c("surface", "heatmap", "contour"), 
            selected = "surface")
checkboxInput("freez", label = "Free z limits?", value = FALSE)

```

Column
--------------------------------------------------------------------------------

### Distribution

```{r, eval=e$vis_distribution}

e$d <- e$sim$sp_N
e$x <- sort(unique(e$d$x))
e$y <- sort(unique(e$d$y))
e$p <- plot_ly(x = e$x, y = e$y)

renderPlotly({
  e$z <- t(xtabs(N ~ x + y, subset = age == input$age & year == input$year, data = e$d))
  if (input$freez) {
    e$r <- range(e$z)
  } else {
    e$r <- range(e$d$N)
  }
  add_trace(e$p, z = e$z, type = input$type, zauto = input$freez, zmin = min(e$d$N), zmax = max(e$d$N)) %>%
    layout(autosize = TRUE,
           scene = list(zaxis = list(title = "N", range = e$r)))
})

```