### 2020-04-09 analysis/paper/response_letter.Rmd
Start response to latest round of peer-review

---
output:
  word_document: default
  md_document: default
---

Daniel E. Duplisea, PhD  
Fisheries and Oceans Canada  
850 Route de la Mer  
Mont‐Joli, Quebec  
G5H 3Z4, Canada

`r format(Sys.time(), '%Y-%m-%d')`

Dear Dr. Duplisea,

Thank you for considering another revision of manuscript PONE-D-19-26904R2,  **"`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks"** by Paul M. Regular, Gregory J. Robertson, Keith P. Lewis, Jonathan Babyn, Brian Healey and Fran Mowbray. We have carefully considered the latest round of suggestions and below we outline how we have addressed each point.

We submit this revised manuscript for your consideration and look forward to your decision.

Sincerely,

Paul Regular  
Fisheries and Oceans Canada  
Northwest Atlantic Fisheries Center  
80 East White Hills, St. John’s, NL  
A1C 5X1, Canada  
E-mail: Paul.Regular@dfo-mpo.gc.ca  
Phone: (709) 772-2067

----------------------------------------------------------------------------------------------------

Reviewer #1: This is my second time reviewing this manuscript and the authors did a great job addressing people's comments. Great job for that. The manuscript is now much clearer and read well, I think.

**We thank the reviewer for the kind words.**

This time, I only have a few minor comments:  
L19 : «there is a limited number of»

**We have removed the 'the' that should not have been in that sentence.**

L103: Maybe it would be a good idea to justify why you are converting to abundance at length.

**Agreed, we have started a new paragraph regarding abundance at length which begins: "In practice, abundance at age is often inferred from length data as it is easier to collect. Abundance at length is therefore simulated from abundance at age using the original von Bertalanffy growth curve"**

L240-242: I think it could be a good idea to show how to exactly do this (maybe obvious to advanced people but not to the general audience). This could be a simple reference to an example on the github repo.

**...**

Table 2: If bias correction was not implemented, I would suggest to change the labeling in the table here and elsewhere (if needed) and replace the “mean” by “median” (if the variable is transformed back to the original scale). If not, I would make clear that you talk about mean in log scale.

L326: I did not see how one can use INLA or RandomFields in Appendix S3. This is a similar comment as the one for line 240-242.

L481-490: the processing time is surprisingly long… it is not an issue in itself but this might refrain some people to using this package… I think some code optimization could help (e.g. vectorize, use matrix operation as much as possible, etc)