### 2021-12-13 tutorial/tutorial.Rmd
Start making learnr tutorial

---
title: "SimSurvey demo"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(SimSurvey)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

### Fisheries-independent surveys

<div style="float: left; width: 50%;">

- Conducted by many RFMOs
- Becoming increasingly important in stock assessment
- Associated with difficult questions:
    - How many fish are in the sea?
    - Where are they located?
    - How should we sample the population?
    - How should we analyze these data?

</div>

<div style="float: right; width: 50%;">

<img src="https://pbs.twimg.com/media/DDRLGEmXUAAXhGf?format=jpg&name=large" width="400"/>
<br>
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
</font>

</div>

## Core functions

### `sim_abundance`

```{r sim-abundance, exercise = TRUE, exercise.eval = FALSE}
set.seed(438)
long <- sim_abundance(ages = 1:20,
                      R = sim_R(log_mean = log(3e+07)),
                      Z = sim_Z(log_mean = log(0.2)))
short <- sim_abundance(ages = 1:6,
                       R = sim_R(log_mean = log(1e+10)),
                       Z = sim_Z(log_mean = log(0.8)))
```