### 2021-12-14 tutorial/tutorial.Rmd
Introduce sim_abundance

---
title: "SimSurvey demo"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(SimSurvey)
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

### Fisheries-independent surveys

<div style="float: left; width: 50%;">

- Conducted by many RFMOs
- Becoming increasingly important in stock assessment
- Associated with difficult questions:
    - How many fish are in the sea?
    - Where are they located?
    - How should we sample the population?
    - How should we analyze these data?

</div>

<div style="float: right; width: 50%;">

<img src="https://pbs.twimg.com/media/DDRLGEmXUAAXhGf?format=jpg&name=large" width="350"/>
<br>
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
</font>

</div>

<br>

<div style="float: left; width: 100%;">

### How can [SimSurvey](https://github.com/PaulRegular/SimSurvey) help?

- A sandbox for simulating of fisheries-independent trawl surveys
  - **Survey:** random or stratified-random
  - **Population:** age-structured and spatially-correlated
- Facilitates tests of the design and analysis of survey data
  - How many sets, lengths, ages are enough?
  - Should the analysis be design or model based?
  - Can models be used to stitch surveys or fill gaps?

</div>

## Simulation steps

1. **Simulate abundance** - `sim_abundance()`
    - Common cohort model

2. **Simulate spatial aggregation** - `sim_distribution()`
    - Includes depth associations and noise correlated across space, years and ages

3. **Simulate survey data** - `sim_survey()`
    - Sample population using stratified random survey
    - These data can be analyzed using `run_strat()` (design-based analysis), or using a package like [sdmTMB](https://github.com/pbs-assess/sdmTMB)

## Simulate abundance

### `sim_abundance()`

- Abundance at age and length is simulated using `sim_abundance()`
- Key arguments:
  - `ages`, `years` - supply vectors of ages and years to simulate across
  - `R`, `Z`, `growth` - supply closures such as `sim_R()`, `sim_Z()`, and `sim_vonB()`

### Closure

- Functions that retain argument values and return functions

```{r closure, exercise = TRUE, exercise.eval = FALSE}
R_fun <- sim_R(log_mean = log(500), log_sd = 0.5)
R_fun
```

### Usage

- Closures minimize the repeated specification of arguments such as ages and years
  - Specify once in `sim_abundance` and the values are used by the functions returned by the closures

```{r usage, exercise = TRUE, exercise.eval = FALSE}
pop <- sim_abundance(ages = 1:6, years = 1:4,
                     R = sim_R(log_mean = log(1000)),
                     Z = sim_Z(log_mean = log(0.8)))
round(pop$N)
```

### Tinker and plot

- For instance, a relatively short and long lived species are simulated below
- Results can be visualized using `plot_surface()`

```{r tinker, exercise = TRUE, exercise.eval = FALSE}
set.seed(438)
long <- sim_abundance(ages = 1:20,
                      R = sim_R(log_mean = log(3e+07)),
                      Z = sim_Z(log_mean = log(0.2)))
short <- sim_abundance(ages = 1:6,
                       R = sim_R(log_mean = log(1e+10)),
                       Z = sim_Z(log_mean = log(0.8)))
plot_surface(short, mat = "N")
```

## Simulate spatial aggregation

### `sim_distribution()`