### 2021-12-15 tutorial/tutorial.Rmd
Introduce the rest of the core functions in the tutorial and start sdmTMB section

---
title: "SimSurvey demo"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(SimSurvey)
library(dplyr)
library(sdmTMB)
library(plotly)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      out.width = "100%", out.height = "450px")
```

## Introduction

### Fisheries-independent surveys

<div style="float: left; width: 50%;">

- Conducted by many RFMOs
- Becoming increasingly important in stock assessment
- Associated with difficult questions:
    - How many fish are in the sea?
    - Where are they located?
    - How should we sample the population?
    - How should we analyze these data?

</div>

<div style="float: right; width: 50%;">

<img src="https://pbs.twimg.com/media/DDRLGEmXUAAXhGf?format=jpg&name=large" width="350"/>
<br>
&nbsp; <a href="https://twitter.com/coastguardcan/status/879410397790515201">Canadian Coast Guard | Twitter</a>
</font>

</div>

<br>

<div style="float: left; width: 100%;">

### How can [SimSurvey](https://github.com/PaulRegular/SimSurvey) help?

- A sandbox for simulating of fisheries-independent trawl surveys
  - **Survey:** random or stratified-random
  - **Population:** age-structured and spatially-correlated
- Facilitates tests of the design and analysis of survey data
  - How many sets, lengths, ages are enough?
  - Should the analysis be design or model based?
  - Can models be used to stitch surveys or fill gaps?

</div>

## Simulation steps

1. **Simulate abundance** - `sim_abundance()`
    - Common cohort model

2. **Simulate spatial aggregation** - `sim_distribution()`
    - Includes depth associations and noise correlated across space, years and ages

3. **Simulate survey data** - `sim_survey()`
    - Sample population using stratified random survey
    - These data can be analyzed using `run_strat()` (design-based analysis), or using a package like [sdmTMB](https://github.com/pbs-assess/sdmTMB)

## Simulate abundance

### `sim_abundance()`

- Abundance at age and length is simulated using `sim_abundance()`
- Key arguments:
  - `ages`, `years` -- supply vectors of ages and years to simulate across
  - `R`, `Z`, `growth` -- supply closures such as `sim_R()`, `sim_Z()`, and `sim_vonB()`

### Closure

- Functions that retain argument values and return functions

```{r closure, exercise = TRUE, exercise.eval = FALSE}
R_fun <- sim_R(log_mean = log(500), log_sd = 0.5)
R_fun
```

### Usage

- Closures minimize the repeated specification of arguments such as ages and years
  - Specify once in `sim_abundance` and the values are used by the functions returned by the closures

```{r usage, exercise = TRUE, exercise.eval = FALSE}
pop <- sim_abundance(ages = 1:6, years = 1:4,
                     R = sim_R(log_mean = log(1000)),
                     Z = sim_Z(log_mean = log(0.8)))
round(pop$N)
```

### Tinker and plot

- For instance, a relatively short and long lived species are simulated below
- Results can be visualized using `plot_surface()`

```{r tinker, exercise = TRUE, exercise.eval = FALSE}
set.seed(438)
long <- sim_abundance(ages = 1:20,
                      R = sim_R(log_mean = log(3e+07)),
                      Z = sim_Z(log_mean = log(0.2)))
short <- sim_abundance(ages = 1:6,
                       R = sim_R(log_mean = log(1e+10)),
                       Z = sim_Z(log_mean = log(0.8)))
plot_surface(short, mat = "N")
```

## Simulate spatial aggregation

### `sim_distribution()`

- `sim_distribution()` is used to distribute a population simulated using `sim_abundance()` throughout a grid
- Key arguments:
  - `sim` -- A list with abundance details like produced by `sim_abundance()`
  - `grid` -- A raster object defining the survey grid, like the one produced by `make_grid()`
  - `ays_covar` -- Closure for simulating age-year-space covariance, like `sim_ays_covar()`
  - `depth_par` -- Closure for defining relationship between abundance and depth, like `sim_parabola()`

### Make a grid

- `make_grid()` sets up a depth stratified survey grid
  - Returns a raster object with four layers: `depth`, `cell`, `division`, and `strat`
  - Number of strata are affected by the number of divisions (`n_div`), horizontal splits (`strat_splits`), the depth gradient (`shelf_depth`, `shelf_width`, `depth_range`), and depth breaks (`strat_breaks`)

```{r grid, exercise = TRUE, exercise.eval = FALSE}
g <- make_grid(res = c(10, 19), n_div = 4, strat_splits = 1, 
               shelf_depth = 500, shelf_width = 0, depth_range = c(0, 1000))
plot_grid(g)
```

### Populating the grid

- The degree of spatial clustering across ages and years can be controled using `sim_ays_covar()` by adjusting the decorrelation range (`range`), and the degree of correlation across age (`phi_year`) and year (`phi_age`) dimensions
- A parabolic depth preference can be controled using `sim_parabola()` where `mu` defines centran tendency and `sigma` the spread

```{r populate-grid, exercise = TRUE, exercise.eval = FALSE}
pop <- sim_abundance(ages = 1:10, years = 1:10) %>%
           sim_distribution(grid = make_grid(res = c(10, 10)),
                            ays_covar = sim_ays_covar(range = 300,
                                                      phi_age = 0.8,
                                                      phi_year = 0.1),
                            depth_par = sim_parabola(mu = 200,
                                                     sigma = 50))
plot_distribution(pop, ages = 1:3, years = 1:3, type = "heatmap")
```

## Simulate survey data

### `sim_survey()`

- Samples from a survey are generated using `sim_survey()`
- Key arguments:
  - `n_sims` -- Number of surveys to simulate over the same population
  - `q` -- Closure, such as `sim_logistic()`, for simulating catchability at age
  - `trawl_dim` -- Trawl width and distance
  - `set_den`, `lengths_cap`, and `ages_cap` -- Sampling protocol for sets, length measurement, and age sampling

### Collect samples

- Conduct a survey with a relatively low and high set density:

```{r survey, exercise = TRUE, exercise.eval = FALSE}
set.seed(438)
pop <- sim_abundance(ages = 1:10, years = 1:5) %>%
  sim_distribution(grid = make_grid(res = c(10, 10)))

a <- pop %>%
  sim_survey(set_den = 1 / 1000)
b <- pop %>% 
  sim_survey(set_den = 5 / 1000)

plot_survey(a, which_year = 2, which_sim = 1)
```

## Analyze

### Design or model-based? That is the question

- Simulated data returned from `sim_survey()` can be analyzed using 
  - a design-based approach (e.g., stratified analysis; built-in `run_strat()` function), or
  - a model-based approach (e.g., geostatistical model; [sdmTMB](https://github.com/pbs-assess/sdmTMB))

### Design-based -- stratified analysis

- `run_strat()` runs a basic stratified analysis and `strat_error()` computes error statistics

```{r strat, exercise = TRUE, exercise.eval = FALSE}
sim <- sim_abundance(ages = 1:5, years = 1:5,
              R = sim_R(log_mean = log(1e+7)),
              growth = sim_vonB(length_group = 1)) %>%
  sim_distribution(grid = make_grid(res = c(10, 10)),
                   ays_covar = sim_ays_covar(sd = 1)) %>%
  sim_survey(n_sims = 1, q = sim_logistic(k = 2, x0 = 3)) %>%
  run_strat() %>%
  strat_error()

sim$total_strat_error_stats
sim$length_strat_error_stats
sim$age_strat_error_stats
```

### Model-based -- [sdmTMB](https://github.com/pbs-assess/sdmTMB)

- sdmTMB is an R package for fast and flexible fitting of spatiotemporal species distribution GLMMs with [TMB](https://github.com/kaskr/adcomp)

- Modeling steps:

  1. **Make an R-INLA "mesh" representing spatial "knots"** - `make_mesh()`
  2. **Fit a spatiotemporal GLMM** - `sdmTMB()`
      - Implements a geostatistical model using TMB, R-INLA, and Gaussian Markov random fields
      - Function is set up similarly to `glm()`, `lmer()`, or `glmmTMB()`
  3. **Inspect the model** - `print()`, `tidy()`, `predict()`, `residuals()`, `sdmTMB_cv()`, `sdmTMB_sim()`, `AIC()`
  4. **Predict on a grid that represents the survey domain** - `predict()`
  5. **Calculate the survey index** - `get_index()`
      - Use `get_index()` to sum the expected density or abundance across the grid and calculate standard errors

### `sdmTMB::sdmTMB()`

```{r sdmTMB, exercise = TRUE, exercise.eval = TRUE}

## Simulate population
set.seed(271)
pop <- sim_abundance(ages = seq(1, 10), years = seq(1, 20)) %>%
  sim_distribution(grid = make_grid(res = c(10, 10), depth_range = c(10, 500)),
                   ays_covar = sim_ays_covar(phi_age = 0.8, phi_year = 0.2),
                   depth_par = sim_parabola(mu = 200, sigma = 30))

## Simulate survey data
survey <- sim_survey(pop, n_sims = 1) %>% run_strat()

## Add coordinates to set details
xy <- as_tibble(survey$grid_xy)
dat <- as_tibble(survey$setdet) %>%
  select(x, y, set, year, N = n, tow_area) %>%
  left_join(., xy, by = c("x", "y")) %>%
  mutate(offset = log(tow_area))

## Make mesh and fit geostatistical model to survey data
mesh <- sdmTMB::make_mesh(dat, xy_cols = c("x", "y"), cutoff = 20)
fit <- sdmTMB(N ~ 0 + as.factor(year) + offset,
              data = dat,
              family = nbinom2(link = "log"), spde = mesh,
              include_spatial = TRUE, time = "year")

## Expand grid data for predictions
grid_dat <- as_tibble(select(pop$grid_xy, x, y, depth)) %>% distinct()
grid_dat <- purrr::map_dfr(sort(unique(dat$year)), ~ bind_cols(grid_dat, year = .))
grid_dat$offset <- mean(dat$offset)

## Predict across full grid and get index
tow_area <- survey$setdet$tow_area[1] # cell and tow area is consistent in the simulation
cell_area <- survey$setdet$cell_area[1]
pred <- predict(fit, newdata = grid_dat, return_tmb_object = TRUE, area = cell_area / tow_area)
index <- get_index(pred)

## Combine indices
true_abund <- tibble(year = unique(dat$year), N = as.numeric(colSums(survey$I))) %>% mutate(type = "True")
strat_abund <- tibble::as_tibble(survey$total_strat) %>%
  mutate(N = total, type = "Design-based") %>%
  select(year, N, type)
index <- index %>%
  mutate(type = "Model-based", N = est) %>%
  bind_rows(strat_abund) %>%
  bind_rows(true_abund)

## Visualize result
index %>% 
  plot_ly(x = ~year, color = ~type, legendgroup = ~type) %>% 
  add_ribbons(ymin = ~lwr, ymax = ~upr, line = list(width = 0), showlegend = FALSE) %>% 
  add_lines(y = ~N)

```